Analyzing Biological Models with the Wasserstein Distance

### 2016-06-03 Paper/Paper.tex
New Paper Updates, Code Updates
implemented explicit finite difference methods for Dirichlet and Neumann BC

\section{Introduction}

@@

\section{One Dimensional Case} \label{sec: OneD}
In one dimension, there exists a closed form solution to computing the Wasserstein distance between two probability density functions. Let $f, g: \mathbb{R} \rightarrow \mathbb{R}$ such that
\begin{equation}
\int_{\mathbb{R}} f \ dx = \int_{\mathbb{R}} g \ dx.
\end{equation}
If we define
\begin{equation}
F(x) = \int_{- \infty}^x f(s) \ ds \ \ \ \ \text{and} \ \ \ \ G(x) = \int_{- \infty}^x g(s) \ ds,
\end{equation}
then the $2$nd Wasserstein distance between $f$ and $g$ is
\begin{equation}
W_2^2(f, g) = {\int_{0}^1 |F^{-1}(y) - G^{-1}(y)|^2 \ dy}.
\label{eq: 1DWD}
\end{equation}
From now on, we will refer to the value of ($\ref{eq: 1DWD}$) as the Wasserstein distance between $f$ and $g$, or simply the WD between $f$ and $g$.
One can think of (\ref{eq: 1DWD}) as moving the $j$-th percentile of $f$ to the $j$-th percentile of $g$ (Figure \ref{fig: 1DWD}).

@@

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale = .49]{1st1DWD.pdf}
  \caption{Discrete Points, WD $\approx$ 1.7475}
  \label{fig: 1Ddiscreet}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale = .5]{2nd1DWD.pdf}
  \caption{Continuous Functions, WD $\approx$ 2.236}
  \label{fig: 1DContinuous}
\end{subfigure}
\caption{Computations of $W_2$ in one dimension}
\label{fig:1DWDCalculations}
\end{figure}

@@

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{WDVSNDX.jpg}
  \caption{Fitting WD to $\frac{N}{\Delta x}$}
  \label{fig:Fitting2DMicroVelocity}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{2DMacroVelocity.jpg}
  \caption{Fitting Log(WD) to $Log(\frac{N}{\Delta x})$}
  \label{fig: Fitting2DMacroVelocity}
\end{subfigure}
\caption{2D Velocities, $T_{final} = 1, \Delta t = .01, N = 10^3, \sigma = \sqrt{2}.$}
\label{fig:test}
\end{figure}

@@

% - Release $Name:  $ -

>>>>

\begin{center}
\textbf{Recent progress:}
\end{center}

\begin{enumerate}
\item Went through the 1D,2D non-bio work. All data files for images in 1D section have been saved to folder ``Data for Paper''. Also fixed wording in section. 
\item For 1D bio work, saved data in folder ``Data for Paper''
\begin{itemize}
\item Put in images of rhoMicro, rhoMacro at time initial and final. 
\item Fitted WD to N, Log(WD) to Log(N) and fitted WD to $\Delta x$, saved data.
\item Computed rhoMicro, rhoMacro velocity, included images, saved data.
\end{itemize}
\item For 2D bio work, 
\begin{itemize}
\item Put in some basic wording
\item Wrote Velocity code, included Macro velocity image and data in folder. 
\item Fixed micro code.
\item Added WD code for 2D (took old simplex method code and call it in 2DStochasticFinal).
\end{itemize}
\textbf{Questions:} 
\begin{itemize}
\item Normalization of Micro
\item How large are our systems for the 2D bio? Even 121 variables takes 4.5 minutes to compute. Might try hungarian algorithm, interior point methods, etc to speed up.
\end{itemize}
\end{enumerate}

\newpage
\section{Introduction}

@@

\section{One Dimensional Case} \label{sec: OneD}
In one dimension, there exists a closed form solution to computing the Wasserstein distance between two probability density functions. Let $f, g: \mathbb{R} \rightarrow \mathbb{R}$ be functions such that
\begin{equation}
\int_{\mathbb{R}} f(x) \ dx = \int_{\mathbb{R}} g(x) \ dx.
\end{equation}
If we define
\begin{equation}
F(x) = \int_{- \infty}^x f(s) \ ds \ \ \ \ \text{and} \ \ \ \ G(x) = \int_{- \infty}^x g(s) \ ds,
\end{equation}
then the $2$nd Wasserstein distance between $f$ and $g$ is
\begin{equation}
W_2^2(f, g) = {\int_{0}^1 |F^{-1}(y) - G^{-1}(y)|^2 \ dy}.
\label{eq: 1DWD}
\end{equation}
From now on, we will refer to the value of ($\ref{eq: 1DWD}$) as the Wasserstein distance between $f$ and $g$, or simply the WD between $f$ and $g$.
One can think of (\ref{eq: 1DWD}) as moving the $j$-th percentile of $f$ to the $j$-th percentile of $g$ (Figure \ref{fig: 1DWD}).

@@

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale = .4]{1st1DWD.pdf}
  \caption{Discrete Points, WD $\approx$ 1.7475}
  \label{fig: 1Ddiscreet}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale = .4]{1DWDNormal.jpg}
  \caption{Continuous Functions, WD $\approx$ 2.06}
  \label{fig: 1DContinuous}
\end{subfigure}
\caption{Computations of WD in one dimension}
\label{fig:1DWDCalculations}
\end{figure}

@@

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{WDVSNDX.jpg}
  \caption{WRONG IMAGE HERE}
  \label{fig:Fitting2DMicroVelocity}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{2DMacroVelocity.jpg}
  \caption{Fitting Log(WD) to $Log(\frac{N}{\Delta x})$}
  \label{fig: Fitting2DMacroVelocity}
\end{subfigure}
\caption{2D Velocities, $T_{final} = 1, \Delta t = .01, N = 10^3, \sigma = \sqrt{2}.$}
\label{fig:test}
\end{figure}

@@
